---
title: "Supplementary_Figure10.rmd"
author: "Michael Noe"
date: "2024-01-26"
output: html_document
---

```{r}
dir.create("../../cfepigenetics_data")
dir.create("../../cfepigenetics_data/cpg_500000_topbot1000")
dir.create("../../cfepigenetics_data/tss_500000_topbot1000")
```


## Process the data

*analysis/code/cpg_500000_topbot1000.sh* -> outputs "../../cfepigenetics_data/cpg_500000_topbot1000"
    Collect information on coverage and size around CpG-islands.
*analysis/code/tss_500000_topbot1000.sh* -> outputs "../../cfepigenetics_data/tss_500000_topbot1000"
    Collect information on coverage and size around TSS-sites.
*analysis/code/cpg_500000_topbot1000_beta.R* -> outputs "../data/cpg_500000_topbot1000_beta.rds"
    Collect information on average beta-values around CpG-islands.
*analysis/code/tss_500000_topbot1000_beta.R* -> outputs "../data/tss_500000_topbot1000_beta.rds"
    Collect information on average beta-values around TSS-sites.
  

## Construct files with metrics around the CpG-islands (1000 methylated and 1000 unmethylated)

```{r}
files <- list.files("../../cfepigenetics_data/cpg_500000_topbot1000", full.names = T)
t <- tibble()
for (file in files) {
  print(basename(file))
  temp <- readRDS(file)
  t <- rbind(t, temp)
}
t[which(t$mean_size == 0),]$mean_size <- NA
t[which(t$mean_size <= 160),]$mean_size <- 160
t[which(t$mean_size >= 176),]$mean_size <- 176
t[which(t$mean_cov == 0),]$mean_cov <- NA
t[which(t$mean_cov <= 400),]$mean_cov <- 400
t[which(t$mean_cov >= 2000),]$mean_cov <- 2000
t <- t %>% 
  group_by(relpos_500, beta) %>%
  summarise(cov = mean(mean_cov, na.rm=T), size = mean(mean_size, na.rm=T))
saveRDS(t, "../data/cpg_500000_topbot1000.rds")
``` 


## Construct files with metrics around the TSS-sites (1000 expressed and 1000 not expressed)

```{r}
files <- list.files("../../cfepigenetics_data/tss_500000_topbot1000", full.names = T)
t <- tibble()
for (file in files) {
  print(basename(file))
  temp <- readRDS(file)
  t <- rbind(t, temp)
}
t[which(t$mean_size == 0),]$mean_size <- NA
t[which(t$mean_size <= 160),]$mean_size <- 160
t[which(t$mean_size >= 176),]$mean_size <- 176
t[which(t$mean_cov == 0),]$mean_cov <- NA
t[which(t$mean_cov <= 400),]$mean_cov <- 400
t[which(t$mean_cov >= 2000),]$mean_cov <- 2000
t$group <- "expressed"
t[which(t$expression <= 5),]$group <- "not expressed"
t <- t %>% 
  group_by(relpos_500, group) %>%
  summarise(cov = mean(mean_cov, na.rm = T), size = mean(mean_size, na.rm = T))
saveRDS(t, "../data/tss_500000_topbot1000.rds")
``` 


