---
title: "Cristiano_Replicate"
output: console
---
Read in Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(devtools)
library(caret)
library(recipes)
library(pROC)
```

Import all data and format metadata
```{r data_read}
delfi<- read_csv("../data/Cristiano-training-set-full.csv") 
meth<-readRDS("../data/carvalho_motifs_ML.rds")

meth<-meth %>% filter(sample %in% delfi$id) #Need to get rid of some LUCAS and D/C healthies that are in here
meth<-meth %>% filter(class_group=="CGPLH"|class_group=="CGPLPA")
meth<-meth %>% dplyr::select(-class) %>% dplyr::rename("type"="class_group")
meth<-meth %>% mutate(type=if_else(type=="CGPLH","healthy","cancer"))
meth_reg <-meth %>% dplyr::select(sample,type,contains("regular"))

data<-inner_join(delfi %>% dplyr::select(-type,-Tumor_Type),meth,by=c("id"="sample"))

#delfi<-inner_join(delfi %>% select(-type,-Tumor_Type),meth %>% select(sample,type),by=c("id"="sample"))
delfi<-delfi %>% dplyr::select(-Tumor_Type)

```

#Now work on the ensembles!
#10 fold -- leave the test set out each time, you have to manually run it for each of 10 times (the 5 cells below)
#fold01 - 09, fold10
```{r}
library(splitTools)
set.seed(1234)
train_folds<-create_folds(meth$type,k=10)

train_IDs<-meth$sample[train_folds$Fold10] ###change this name 10 times (Fold01, Fold02, Fold03, ...)
test_IDs<-meth$sample[-train_folds$Fold10] ###change this name 10 times (Fold01, Fold02, Fold03, ...)
name<-"../data/fold10_ens.csv" ###change this name 10 times
```

#train meth GBM 5x1 on train set and save CV
#train DELFI LR 5x1 on train set and save CV
#Of note, we are spiking in the Cristiano cancers of other types as extra training data
#use CV preds to train ensemble

#predict DELFI and Meth on test set
#predict ensemble on test set

```{r}
set.seed(1234)
meth_data_train<-meth %>% filter(sample %in% train_IDs)
meth_data_test<-meth %>% filter(sample %in% test_IDs)

recipe_meth <- recipe(type ~ ., data=meth_data_train) %>%
    update_role(sample, new_role = "ID") %>%
    step_corr(all_predictors(), threshold=0.95) %>%
    step_nzv(all_predictors())
set.seed(1234)
ctrl_meth <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 1,
                     verboseIter = TRUE,
                     savePredictions="final",
                     classProbs=TRUE,
                     index=createMultiFolds(meth_data_train$type, 5, 1),
                     summaryFunction = twoClassSummary)

set.seed(1234)
model_meth <- caret::train(recipe_meth,
                      data = meth_data_train,
                      method = "gbm",
                      trControl = ctrl_meth)

##################
set.seed(1234)
meth_reg_data_train<-meth_reg %>% filter(sample %in% train_IDs)
meth_reg_data_test<-meth_reg %>% filter(sample %in% test_IDs)

recipe_meth_reg <- recipe(type ~ ., data=meth_reg_data_train) %>%
    update_role(sample, new_role = "ID") %>%
    step_corr(all_predictors(), threshold=0.95) %>%
    step_nzv(all_predictors())
set.seed(1234)
ctrl_meth_reg <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 1,
                     verboseIter = TRUE,
                     savePredictions="final",
                     classProbs=TRUE,
                     index=createMultiFolds(meth_reg_data_train$type, 5, 1),
                     summaryFunction = twoClassSummary)

set.seed(1234)
model_meth_reg <- caret::train(recipe_meth_reg,
                      data = meth_reg_data_train,
                      method = "gbm",
                      trControl = ctrl_meth_reg)
#############
delfi_data_train<-delfi %>% filter(!id %in% test_IDs)
delfi_data_test<-delfi %>% filter(id %in% test_IDs)

recipe_delfi <- recipe(type ~ ., data=delfi_data_train) %>%
    update_role(id, new_role = "ID") %>%
    step_rm(starts_with("ratio"),starts_with("short")) %>%
    step_pca(starts_with("cov"), prefix = "cov_pc_",  threshold=0.90)     %>%
    step_corr(all_predictors(), threshold=0.95) %>%
    step_nzv(all_predictors())

glmnetGrid <- expand.grid(
    alpha = 1,
    lambda = 10^seq(-5, -1, length.out = 100))
#### Train models
set.seed(1234)
ctrl_delfi <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 1,
                     verboseIter = TRUE,
                     savePredictions="final",
                     classProbs=TRUE,
                     index=createMultiFolds(delfi_data_train$type, 5, 1),
                     summaryFunction = twoClassSummary)

set.seed(1234)
model_delfi <- caret::train(recipe_delfi,
                          data = delfi_data_train,
                          method = "glmnet",
                          tuneGrid = glmnetGrid,
                          trControl = ctrl_delfi)
##########
pred.all <- model_meth$pred
pred.all <- pred.all %>% dplyr::group_by(rowIndex) %>% dplyr::summarize(score.meth = mean(cancer))
all_data <- meth_data_train %>% dplyr::mutate(rowIndex = 1:n())
labels <- all_data %>% dplyr::select(type,sample,rowIndex)
preds_all <- dplyr::inner_join(labels, pred.all, by="rowIndex")

pred.all <- model_meth_reg$pred
pred.all <- pred.all %>% dplyr::group_by(rowIndex) %>% dplyr::summarize(score.meth_reg = mean(cancer))
all_data <- meth_reg_data_train %>% dplyr::mutate(rowIndex = 1:n())
labels <- all_data %>% dplyr::select(type,sample,rowIndex)
preds_all_reg <- dplyr::inner_join(labels, pred.all, by="rowIndex")

pred.all <- model_delfi$pred
pred.all <- pred.all %>% dplyr::group_by(rowIndex) %>% dplyr::summarize(score.delfi = mean(cancer))
all_data <- delfi_data_train %>% dplyr::mutate(rowIndex = 1:n())
labels <- all_data %>% dplyr::select(id,rowIndex)
preds <- dplyr::inner_join(labels, pred.all, by="rowIndex")

preds_CV <- inner_join(preds_all,preds,by=c("sample"="id"))
#colnames(preds_CV) <- sub("rowIndex.x", "rowIndex", colnames(preds_CV))
preds_CV <- inner_join(preds_CV,preds_all_reg,by=c("sample", "type"))



recipe_ens <- recipe(type ~ score.meth_reg + score.delfi, data=preds_CV)

set.seed(1234)
ctrl_ens <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 1,
                     verboseIter = TRUE,
                     savePredictions="final",
                     classProbs=TRUE,
                     index=createMultiFolds(preds_CV$type, 5, 1),
                     summaryFunction = twoClassSummary)




set.seed(1234)
model_ens <- caret::train(recipe_ens,
                          data = preds_CV,
                          method = "glmnet",
                          tuneGrid = glmnetGrid,
                          trControl = ctrl_ens)
#######
#Delfi Preds
delfi_preds<-predict(model_delfi,delfi_data_test,type="prob")$cancer
meth_preds<-predict(model_meth,meth_data_test,type="prob")$cancer
meth_reg_preds<-predict(model_meth_reg,meth_reg_data_test,type="prob")$cancer

preds<-cbind(meth_data_test %>% dplyr::select(sample,type),meth_preds)
preds_reg<-cbind(meth_reg_data_test %>% dplyr::select(sample,type),meth_reg_preds)
preds2<-cbind(delfi_data_test %>% dplyr::select(id),delfi_preds)
preds<-inner_join(preds,preds2,by=c("sample"="id"))
preds<-inner_join(preds,preds_reg,by=c("sample", "type"))
preds<-preds %>% dplyr::rename(score.delfi=delfi_preds) %>% dplyr::rename(score.meth=meth_preds) %>% dplyr::rename(score.meth_reg=meth_reg_preds)
ens_preds<-predict(model_ens,preds,type="prob")$cancer

preds<-cbind(preds,ens_preds)

write.csv(preds,name)

```



After you have generated all 10 folds above!!!!
Read all the results
```{r}
data <- list.files(path="../data",pattern="fold*", full.names = TRUE) %>% 
  lapply(read_csv) %>% 
  bind_rows 

write.csv(data,"../data/All_Scores.csv")
```






